Secure Composition of Robust and Optimising Compilers

### 2023-06-19 popl/main.tex
Add more defs

% Fixing: Too many math alphabets used in version normal.
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[acmsmall,review,screen,dvipsnames]{acmart}

@@

@@

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}
\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}
\subsection{Optimizing Compilers}\label{subsec:cs:opts}
\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}

>>>>

% Fixing: Too many math alphabets used in version normal.
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[utf8,acmsmall,review,screen,dvipsnames]{acmart}

@@

% https://tex.stackexchange.com/questions/648845/sans-serif-uppercase-greek-no-longer-showing-in-acmart
\DeclareMathAlphabet{\mathsf}{OT1}{LibertinusSans-LF}{m}{n}
\SetMathAlphabet{\mathsf}{bold}{OT1}{LibertinusSans-LF}{bx}{n}

@@

\subsection{$\src{L_{\tmssafe}}$: A Temporal but Not Spatial Memory Safe Language}\label{subsec:ltms}
\begin{gather*}
  \begin{aligned}
    \mi{(Pre\text{--}Types)}~\src{\type[_{e}]}&\bnfdef\src{\natt} \mid \src{\ptrn} \hspace{0.5cm}
    \mi{(Pointer\ Qualifiers)}~\src{\ptrqual}\bnfdef\ \src{\fullq} \mid \src{\halfq} \\
    \mi{(Type~Environment)}~\src{\typenv}&\bnfdef\ \src{\hole{\cdot}} \mid \hastype{\src{x}}{\src{\type}},\src{\typenv} \hspace{0.5cm}
    \mi{(Types)}~\src{\type}\bnfdef\src{\type[_{e}]} \mid \src{\type[_{e}]\to\bot} \mid \src{\type[^{(1)}_{e}]\to\type[^{(2)}_{e}]}
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\tcheck{\src{\typenv}}{\src{\expr}}{\src{\type}}}{,,Under environment $\src{\typenv}$, expression $\src{\expr}$ has type $\src{\type}$.''}
  %
  \typerule{t-$\src{x}$}{
    \src{\typenv}(\src{x})=\src{\type} &
    \notowned{\src{\typenv}}
  }{
    \tcheck{\src{\typenv}}{\src{x}}{\src{\type}}
  }{t-x}
  %
  \typerule{t-$\src{get}$}{
    \tcheck{\src{\typenv[_2]}}{\src{x}}{\src{\wptr}} &
    \tcheck{\src{\typenv[_1]}}{\src{\expr}}{\src{\natt}} &
  }{
    \tcheck{\typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}}{\src{\lget{x}{\expr}}}{\src{\natt}}
  }{t-get}
  %
  \typerule{t-$\src{let}$}{
    \tcheck{\src{\typenv[_1]}}{\src{\expr[_1]}}{\src{\type[_e]}} &
    \tcheck{\hastype{\src{x}}{\src{\type[_e]}},\src{\typenv[_2]}}{\src{\expr[_2]}}{\src{\type}}
  }{
    \tcheck{\typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}}{\src{\llet{x}{\expr[_1]}{\expr[_2]}}}{\src{\type}}
  }{t-let}
  %
  \typerule{t-$\src{delete}$}{
    \src{\typenv}(\src{x})=\src{\ptr} &
    \notowned{\src{\typenv}\setminus\left\{\hastype{\src{x}}{\src{\ptr}}\right\}}
  }{
    \tcheck{\src{\typenv}}{\src{\ldelete{x}}}{\src{\natt}}
  }{t-delete}
  %
  \typerule{t-$\src{abort}$}{
    \notowned{\src{\typenv}}
  }{
    \tcheck{\src{\typenv}}{\src{\labort}}{\src{\type}}
  }{t-abort}
  %
  \typerule{t-$\src{call}$}{
    \inttype{\src{\type[^{(1)}_e]}} &
    \inttype{\src{\type[^{(2)}_e]}} \\
    \tcheck{\src{\typenv}}{\src{foo}}{\src{\type[^{(1)}_e]\to\type[^{(2)}_e]}} &
    \tcheck{\src{\typenv}}{\src{\expr}}{\src{\type[^{(1)}_e]}}
  }{
    \tcheck{\src{\typenv}}{\src{\lcall{foo}{\expr}}}{\src{\type[^{(2)}_e]}}
  }{t-call}
  %
  \typerule{t-$\src{return}$}{
    \inttype{\src{\type[_e]}} &
    \tcheck{\src{\typenv}}{\src{\expr}}{\src{\type[_e]}}
  }{
    \tcheck{\src{\typenv}}{\src{\lreturn{\expr}}}{\src{\type[_e]\to\bot}}
  }{t-return}
\end{center}

$\src{L_{\tmssafe}}$ is a statically typed language where each well-typed term is temporal memory safe.
Most parts of the type system are entirely standard, so only the interesting bits are shown here.
To enforce this, the language marks pointers with respect to their ownership, where $\src{\fullq}$ means {\em owned} and $\src{\halfq}$ means {\em not owned}.
An owned pointer can only be used by a deallocation (\Cref{tr:t-delete}), whereas reads and writes must be done on non-owned pointers.
One can convert owned to non-owned pointers implicitly through context splitting $\typenvsplit{\src{\typenv[_{1}]}}{\src{\typenv[_{2}]}}$.\MK{list some rules and explain it a bit}
Syntactic constraints together with the type system ensure that non-owned pointers never ,,live'' longer than owned pointers.
For example, in an expression $\src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}}$, if $y$ is a $\src{\ptr}$ in $\src{\expr[_{1}]}$, it cannot appear at all in $\src{\expr[_{2}]}$, but it can if it were of type $\src{\wptr}$ (\Cref{tr:t-let}).
This mechanism also prevents terms like $\src{\lget{x}{\ldelete{x}}}$ from typechecking (\Cref{tr:t-get}).
At leafs of the type derivation tree, the typing environment must not contain owned pointers, which is ensured with the $\notowned{\src{\typenv}}$ judgement (\Cref{tr:t-abort,tr:t-x}).
Lastly, calling or returning requires passed values to have an interface type ($\inttype{\src{\type[_{e}]}}$), which are just $\src{\natt}$ (\Cref{tr:t-call,tr:t-return}).
To ensure that no pointers live after a return, the type of a return is that of a continuation, which is only permitted to propagate towards the end of a term.
For example, we cannot bind $\src{\lreturn{e}}$ to an identifier in the typecheck of $\src{let}$, because the expression whose result is written into the binder has to be a pre--type, which the type of $\src{\lreturn{e}}$ is not.

\MKin{add filter $\theta$ projecting to specification events}

\begin{theorem}[$\src{L}$-programs are \gls{tms}]\label{thm:wt:tms}
  If $\progstepto{\src{\prog{\library_{\ctx}}{\library_{\comp}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\trace}}$, then $\montmssafe{\filterLtms{\src{\trace}}}$.\Coqed
\end{theorem}

\subsection{$\trg{L}$: An Unsafe Language}\label{subsec:ltms}

\MKin{add definitions and description; some syntax + dynamics}

\subsection{Robust Temporal Memory Safety Preservation}\label{subsec:cs:tms}

\MKin{compiler $\src{L}\to\trg{L}$}
\MKin{x-lang relations}
\MKin{backtranslation}
\MKin{rtp of tms}

\subsection{Robust (Spatial) Memory Safety Preservation}\label{subsec:cs:ms}

\MKin{compiler $\trg{L}\to\irl{L_{MS}}$}
\MKin{backward sim}
\MKin{rtp of sms}

\MKin{composition: rtp of ms}

\subsection{Optimizing Compilers}\label{subsec:cs:opts}

\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{dce}}
\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{cf}}

\MKin{backward sims}
\MKin{rtp of ms}
\MKin{composition: rtp of ms, order independent}

\subsection{$\obj{L_{\scct}}$: A Language with Access to a Model Specific Register}\label{subsec:ltms}

\MKin{add definitions and description; some syntax + dynamics}

\subsection{Robust Strict Cryptographic Constant Time Preservation}\label{subsec:cs:scct}

\MKin{compiler $\irl{L_{MS}}\to\obj{L_{\scct}}$}
\MKin{x-lang relations}
\MKin{backtranslation}
\MKin{rtp of scct}

\MKin{composition: rtp of everything}
